Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow

### 2009-05-03 paper/paper.tex
Updated paper.

%\begin{frontmatter}
\newcommand{\omitt}[1]{}
\begin{document}
\title{Extending the Advanced Forensic Format to accommodate Multiple
  Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow}
\author{Blinded for review}
%\author{M. I. Cohen\footnote{scudette@gmail.com}, Simson Garfinkel, and Bradley Schatz}
%\author{M. I. Cohen}
%\address{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
%\baselineskip=20pt %Sets line spacing to 1 unit
%\end{frontmatter}
%\thanks{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
\maketitle

\begin{abstract}
Forensic analysis requires the acquisition and management of many
different types of evidence, including individual disk drives, RAID
sets, network packets, memory images, and extracted files. Often the
same evidence is reviewed by several different tools or examiners in
different locations. We propose a backwards-compatible evolutionary
redesign of the Advanced Forensic Format---an open, extensible file
format for storing and sharing of evidence, arbitrary case related
information and analysis results among different tools. The new
specification was designed to be simple to implement, allowing the use
of the well supported Zip File format specifications for bit level
file access.
\end{abstract}

\section{Introduction}
Unlike other kinds of evidence, digital evidence can be instantly lost
or corrupted if technical measures are not immediately taken to
preserve its existence and integrity. As a result, preserving digital
evidence is an important part of most digital
investigations\cite{carrier:event-based}.

\subsection{Prior Work}
In recent years there has been a steady and growing interest in the
actual file formats and containers used to store this digital
evidence. Early practitioners created exact bit-for-bit copies
(``mirror disks'') or employed proprietary software systems
(e.g. \cite{safeback,ilook,encase}) for making and authenticating
``images'' of digital evidence. PyFlag\cite{pyflag} introduced a
``seekable gzip'' format that allowed disk images to be stored in a
form that was compressed but allowed random-access to evidence data
necessary for forensic analysis. The \emph{sgzip} format does not
store any metadata at all.

The Advanced Forensic Format (AFF) expanded on this idea with a
forensic file format that allowed both data and arbitrary metadata to
be stored in a single digital archive\cite{garfinkel:aff}. The EWF
file format also provides for a small number of predefined metadata
fields to be stored within the file format\cite{libewf}.

Both of these approaches are limited in their expressiveness, by
nature of their (property, value) data model. The subject of the
properties is implicitly the single image contained within the
evidence container. Clearly an implicit subject metadata model is
inadequate for storing and managing multiple evidence sources
simultaneously.

The Sealed Digital Evidence Bags architecture of Schatz proposed an
approach to digital evidence containers that facilitated composition
of evidence and arbitrary evidence related information, through a
simple data model and globally unique referencing
scheme\cite{schatz:sdeb}.

Recently the open source AFF implementation library (AFFLIB) was
expanded to support both encryption and digital signatures of digital
evidence\cite{garfinkel:affcrypto}. Clearly, the implementation of
encryption and authentication is critical for forensic file formats.

\subsection{This paper}
The AFF4 metadata model described in this work, is based on (subject,
property, value) statements. An explicitly defined subject (identified
by an AFF4 URN) allows one to not only associate traditional EWF style
metadata with a contained image, but to also define new information
entities representing both real world and virtual entities related to
the matter. For example, an AFF4 compliant hard disk acquisition tool
should, in addition to storing an image, store entities representing
and describing the tool itself, the operator of the tool, the user
supplied parameters supplied to the tool, and the drive being imaged.

An important advance of this work is the introduction of storage transformation
functions to the forensic storage container. Prior works simply focused
on forensically sound storage of bitstreams, leaving the necessary activities
of translating low level storage into higher level abstractions at the 
aggregate block (ie. RAID), volume, and filesystem layers in the domain of 
analysis tools, as transiently constructed artifacts. AFF4, through its introduction 
of Map Streams, enables artefacts at all these layers to be stored at zero cost within
the forensic container, which is an important enabler for interoperable 
forensic tools.

This paper extends the previous work by adapting AFF to the multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. We do this in a manner that
is at once upwards compatible with AFF while adopting the AFF
bit-level specification so that AFF files can be processed using
standard non-forensic tools. We call the new system AFF4, and use the
phrase AFF1 to refer to the legacy system developed by
Garfinkel.\footnote{Although Garfinkel never changed the AFF bit-level
specification, Garfinkel released AFFLIB implementations with major
version numbers 1, 2 and 3. We therefore call our system AFF4 to avoid
confusion.}

\begin{figure}
\fbox{
\begin{minipage}{.95\linewidth}
\begin{itemize}
\item The ability to store evidence from one or more physical
  devices.
\item Schema allowing evidence files to contain memory
  images, network packet intercepts, Netflow data, and extracted
  files.

\item Direct encoding of arbitrary mapping transformations of one or
more sources to allow multiple logical views. This allows 
RAID volumes to be automatically reassembled from evidence files made
of the physical RAID devices. Maps can be created using Sleuthkit and
stored in the evidence file, thereafter allowing arbitrary programs to directly
read individual files stored in a disk image.

\item A revised compression system that allows individual chunks
  within a 16MB segment to be decompressed without decompressing the
  entire segment.

\item Support for Global Distributed Evidence Management.
\end{itemize}
\end{minipage}}
\caption{Specific design goals for the next generation forensic file
  format.}
\end{figure}
\section{The Need for an Improved Forensic Format}

AFF1's flexibility came from a data model of forensic data and metadata
stored as arbitrary name/value pairs called \emph{segments}. For
example, the first 16MB of a disk image is stored in a segment called
\texttt{page0}, the second 16MB in a segment called \texttt{page1},
\emph{etc.} Because of this flexibility, it was relatively easy for
Garfinkel to extend AFF1 to support encryption, digital signatures, and
the storage of new kinds of metadata such as chain-of-custody
information\cite{garfinkel:affcrypto}.

\subsection{AFF Limitations}
After imaging literally thousands of devices using Garfinkel's AFF1
tools and adding support for AFF1 to numerous open source programs, we
observed numerous problems in the underlying standard
and Garfinkel's AFFLIB implementation:

@@

@@

\item Although the AFF1 specification calls for a ``table of contents'' similar
  to the Zip\cite{zip-format} ``central directory'' that is stored at
  the end of AFF files, Garfinkel never implemented this directory in
  the publicly released AFF1 implementation, AFFLIB. As a result,
  every header of every segment in an AFF file needs to be read when a
  file is opened. In practice this can take 10--30 seconds the first
  time a large AFF file is opened. (Once the file is opened, the
  sectors corresponding to the segment headers are stored in the
  computer's disk cache.)

@@

\section{Global Distributed Evidence Management}
While Garfinkel designed AFF for use on a single machine that could
both image evidence and perform analysis, the main use of AFF in
practice has been in distributed environments in which imaging and
analysis takes place in multiple locations and is performed by
multiple individuals. This is a state of affairs which mirrors common
practice in the field.

@@

Managing evidence in a globally distributed system requires the use of
globally unique identifiers. AFF1 assigns
each piece of evidence a unique 128-bit identifier called a GID but did
not make it clear when this identifier should be changed and when it
should remain the same.

Consider the typical usage scenario depicted in Figure \ref{usage}, of
a volume containing a disk image. This volume is distributed to two
independent analysts, Alice and Bob. Alice may find and extract
individual files, while Bob may correlate information in the evidence
file with other data that is available on departmental
servers. Although in some environments Alice and Bob may be able to
work on a shared file that is located on a server, in other
environments there will not be sufficient bandwidth or
connectivity. Instead, each analyst will be required to store the
information in their own evidence file; these files will then be
recombined at a later point in time.

@@

We have also adopted AFFLIB to implement the new schema and file
format, which will allow existing AFF-aware programs to access images
written in either the old or the new format without modification. We
are extending the AFFLIB API to support the new information model.

We call the result of our work effort AFF4.

@@

This section discusses the AFF4 terminology. The AFF4 design is object
oriented, in that a few generic objects are presented with externaly
accessible behaviour. We discuss a number of implementations of these
high level concepts and show how these can be put together in common
usage cases.

@@

\label{definitions}
\item \emph{A segment} is a single unit of data written to a volume. AFF4
  segments have a \emph{segment name} provided by their URN, a
  \emph{segment timestamp} in GMT, and the \emph{segment
contents}. Segments are suitable for storing small quantities of data,
and still present a stream interface.

\item \emph{A stream} is an AFF Object which provides the ability to
seek and read random data. Stream objects implement abstracted
storage, but must provide clients the stream like interface. For
example, we discuss the {\em Image stream} used to store large images,
the {\em Map stream} used to create transformations and the {\em
Encrypted stream} used to provide encryption.

@@

\section{Metadata and the Universal Resolver}
\label{resolver}
One of the fundamental tasks of forensics is in ensuring the
provenance of any particular piece of digital evidence, for example by
employing a chain of custody process. To assist in this, acquisition
methodologies emphasise the process of individuating evidence, and
storing it in a manner that is easily identifiable.

@@

In this way we can identify uniquely any evidence without the chance
for name collisions. Once a universal identification regime is
created, it is now possible to collect metadata about each object. The
AFF4 model treats metadata as an abstract concept which may exist
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in triple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. Using this system we are able to store
arbitrary attributes about any object in the AFF4 universe. Additionally,
as these statements are universally scoped, they may be stored anywhere.

The AFF4 design extends beyond the management of a single volume,
stream or image to a universal system for managing data of many
types. This necessarily means that a single running instance is
generally unable to have visibility of the entire AFF4 universe. For
example, if a volume is opened which contains a Map Stream targeting a
stream stored in a different volume, its not generally possible to
tell in which volume the stream exists, and where that volume is
actually stored.

To provide this global visibility of metadata we define a central
metadata management entity, named the {\em Universal Resolver}. The
Universal Resolver contains all the metadata about the AFF4 universe,
that is to say it is able to resolve queries about any attribute about
any URN in the universe.

@@

To this end we define a way for serializing metadata statements (or
triples) into a standard format which implementations can load into
their respective resolvers when parsing the volume. Triples can be
stored in segments having a URN ending with ``{\em properties}''. The
AFFLIB implementation loads these segments automatically into the
Universal Resolver.

@@

When the volume is loaded, the AFFLIB implementation automatically
loads any properties files and populates its Universal Resolver with
the information visibile to it. AFFLIB provides a mechanism to use an
external resolver stored on a MySQL database to provide a Universal
Resolver that shares information between different instances on the
same network.

@@

\section{Volumes}
The volume object is responsible for providing storage for
segments. Segments are stored and retrieved using their URNs. We
describe two different implementations of volume objects, namely the
{\em Directory Volume} and the {\em ZipFile Volume}. It is possible to
convert from one implementation to another easily, without effecting
any external references. This is because external references are to
the Volume URN itself, so its safe to convert from one representation
to another. In this way its possible for example, to unzip a Zip64
volume into a directory and vice versa.

@@

\begin{itemize}
\item There is already wide support for the Zip and Zip64 formats. By
  migrating to these formats, we can take advantage of the rich number
  of user and developer tools already available. The volume may be
  inspected using any number of commercial zip application
  (e.g. Windows Explorer natively supports Zip files as can be seen in
  Figure~\ref{explorer}).

@@

  \caption{The basic structure of a Zip archive. Also shows is how new
  archive members are added to an existing Zip File. The Central
  Directory gets overwritten by the new member, and a new Central
  Directory is written on the end.  }

@@

\item Although Zip64 also defines encryption and authentication
extensions, we do not use them due to the restrictions imposed on
their use and because they lack the functionality that is important
for a forensic user. Instead, we use AFF1's digital signature
facilities for integrity and non-repudiation, and we introduce a new
stream based encryption scheme for ensuring data privacy
(Section~\ref{crypted_stream}).

@@

\subsection{Implementing Zip64 in AFFLIB}
Although there are numerous Zip implementations available
today, we have created our own implementation using a combination of
available open source technology and our own novel contributions. We
are in the process of integrating this code into the AFFLIB library
and performing extensive regression testing.

There are many reasons to develop our own Zip64 implementation for
AFF4:

@@

Storage for the data is done by using multiple data segments stored on
various volumes. Data segment URNs are derived by appending an 8
digit, zero padded decimal integer representation of an incrementing
id to the stream URN (e.g. ``urn:aff4:83a3d6db-85d5/00000032''). Each
data segment is called a bevy and stores a number of compressed chunks
back to back.

@@

Image streams specify the \texttt{chunk\_size} attribute, as the
number of image bytes each chunk contains (typically chunk sizes are
32kb). Also specified is the \texttt{chunks\_per\_segment} attribute
which specifies how many chunks are stored in each bevy. Each chunk is
compressed individually using the zlib compress algorithm. This
general structure of storing chunks within larger segments is similar
to the technique used by the Expert Witness file format (EWF) used by
EnCase\cite{encase-3.0} and implemented by the open source
libewf\cite{libewf} package. The advantages of keeping smaller sized
chunks is the better match between requested size and the minimum size
required for decompression. This means that less data is needed to be
decompressed unnecessarily.

@@

The mapping function is described in a segment named by appending
``/map'' to the stream URN. The segment data consists of a series of
lines, each containing a stream offset, followed by a comma, and a
target offset. Offsets are encoded using decimal notation, and are
followed by a target URN.

@@

The Encrypted Stream may contain any data at all, including disk
images, network captures or memory images. It is useful however, to
store an entire AFF4 volume within the Encrypted stream. This provides
block level encryption for the contained AFF4 volume (which might
contain arbitrary streams). It is recommended that Encrypted Streams
containing an AFF4 volume set an attribute
\emph{content-type=application/x-aff-volume} on the container stream,
to allow implementations to automatically recognize volumes. This
approach is illustrated in Figure~\ref{crypted_fif}.

@@

The result is that a number of AFF4 volumes are used as {\em Container
Volumes} to provide storage for Encrypted Streams. The main {\em
Embedded Volume}, which actually contains data is stored within the
Encrypted Stream, effectively distributed throughout the container
volumes. Note that the outer Volume may contain several Encrypted
Streams and therefore contain multiple AFF4 Encrypted 
Volumes. Container Volumes may contain non encrypted streams as well,
and may implement different encryption schemes and keys for each
Encrypted scheme. This effectively allows arbitrary access policies to
be implemented as only volumes which can be accessed can be merged.

\subsection{Encryption Schemes}
Our implementation provides a number of encryption schemes.

\subsubsection{Null Encryption}
This scheme ({\em scheme=null}) specifies no encryption at all. This
may be useful for testing an implementation, but obviously provides
no real security.

\subsubsection{AES-SHA-PSK}
This scheme ({\em scheme=aes-sha-psk}) uses AES for block level
encryption, with the following SHA based ESSIV scheme, and Pre-Shared
Key (passphrase).

When creating a new encryption stream, a master key is generated using
the first 128 bits of a SHA1 hash of the Pre-Shared-Key appended to a
64 bit random salt:
\begin{eqnarray}
Key_{Master} = \left | SHA1(salt + PSK) \right | _{128}
\end{eqnarray}

For each chunk, the chunk IV is obtained by taking the first 128 bits
of the SHA1 hash of the chunk id encoded as a 32 bit little endian
integer, followed by the master key:
\begin{eqnarray}
IV = \left | SHA1(chunk\_id + Key_{Master}) \right | _{128}
\end{eqnarray}

The block is then encrypted or decrypted using AES in CBC mode with
this IV.

The salt is specified as a stream attribute named {\em aff4:salt} in
base64 encoding. Implementations are free to specify any API for
passing the Pre-Shared Key, for example the PSK can be specified in a
configuration file, typed in at the terminal, passed in using process
environment variables, or encoded within the AFF4 URL itself.

@@

\subsection{In-place encryption of AFF4 volumes}
With the large size of a typical forensic image it is often difficult
to have to create a second copy of a volume which needs to be
encrypted. It is usually more convenient to encrypt the file in place,
overwriting the plain text data with the equivalent encrypted data.

>>>>

%\begin{frontmatter}
\newcommand{\omitt}[1]{}
\begin{document}
\title{Extending the Advanced Forensic Format to accommodate Multiple
  Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow}
\author{M. I. Cohen\footnote{scudette@gmail.com}, Simson Garfinkel, and Bradley Schatz}
%\author{M. I. Cohen}
%\address{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
%\baselineskip=20pt %Sets line spacing to 1 unit
%\end{frontmatter}
%\thanks{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
\maketitle

\begin{abstract}
Forensic analysis requires the acquisition and management of many
different types of evidence, including individual disk drives, RAID
sets, network packets, memory images, and extracted files. Often the
same evidence is reviewed by several different tools or examiners in
different locations. We propose a backwards-compatible redesign of the
Advanced Forensic Format---an open, extensible file format for storing
and sharing of evidence, arbitrary case related information and
analysis results among different tools. The new specification, termed
AFF4, was designed to be simple to implement, allowing the use of the
well supported Zip File format specifications for bit level file
access.
\end{abstract}

\section{Introduction}
Storing and managing digital evidence is becoming increasingly more
difficult, as the volume and size of digital evidence increases. The
type of what consitutes evidence has also evolved to include data
other than disk images, such as memory images, network images and
regular files. Preserving digital evidence is an important part of
most digital investigations\cite{carrier:event-based}, and managing
the evidence in a distributed organization is now emerging as a
critical requirement.

This paper presents a framework for managing and storing digital
evidence. We first examine existing evidence management file formats
and outline their strengths and limitations. We then explain how the
proposed Advanced Forensics Format (AFF4) framework extends these
efforts into a universal evidence management system. The detailed
description of the AFF4 proposal is then followed by concerete real
world use cases.

\subsection{Prior Work}
In recent years there has been a steady and growing interest in the
actual file formats and containers used to store digital
evidence. Early practitioners created exact bit-for-bit copies
(commonly referred to as ``dd images''). More recently, proprietary
software systems for making and authenticating ``images'' of digital
evidence are commonly used
(e.g. \cite{safeback,ilook,encase}). PyFlag\cite{pyflag} introduced a
``seekable gzip'' format that allowed disk images to be stored in a
form that was compressed but allowed random-access to evidence data
necessary for forensic analysis. The \emph{sgzip} format does not
store any metadata at all.

The Advanced Forensic Format (AFF) expanded on this idea with a
forensic file format that allowed both data and arbitrary metadata to
be stored in a single digital archive\cite{garfinkel:aff}.

The Eye Witness Forensic (EWF) file format was originally developed
for Encase\cite{encase}, but then adopted by other vendors
\cite{libewf}. The EWF file format also compresses the image into 32kb
chunks which are stored back to back in groupings inside the file. The
format contains multiple tables containing relative indexes to these
compressed chunks which are also stored within the volume. EWF volumes
have a maximum size of 2Gb and therefore usually split an image across
many volumes. EWF provides for a small number of predefined metadata
fields to be stored within the file format.

Both the AFF and EWF file formats are designed to store a single
image, and any metadata implicitely refers to that image. Therefore,
these formats in themselves have a limited metadata model which is
unable to exist outside the context of the image itself.

%Both Of These Approaches Are Limited In Their Expressiveness, By
%Nature Of Their (Property, Value) Data Model. The Subject Of The
%Properties Is Implicitly The Single Image Contained Within The
%Evidence Container. Clearly An Implicit Subject Metadata Model Is
%Inadequate For Storing And Managing Multiple Evidence Sources
%Simultaneously.

Schatz proposed a Sealed Digital Evidence Bags architecture,
facilitating composition of evidence and arbitrary evidence related
information, through a simple data model and globally unique
referencing scheme\cite{schatz:sdeb}.

%Recently the open source AFF implementation library (AFFLIB) was
%expanded to support both encryption and digital signatures of digital
%evidence\cite{garfinkel:affcrypto}. Clearly, the implementation of
%encryption and authentication is critical for forensic file formats.

\subsection{This paper}
The AFF4 metadata model described in this work, goes beyond merely
specifying metadata as an addition to the forensic evidence
itself. Metadata forms an essential part of the file format, with the
same metadata system storing both essential information required to
read the image itself, as well as non-essential metadata. For example,
an AFF4 compliant hard disk acquisition tool should, in addition to
storing an image, store entities representing and describing the tool
itself, the operator of the tool, the user supplied parameters
supplied to the tool, and the drive being imaged.

An important advance of this work is the introduction of storage
transformation functions to the forensic storage container. Prior
works simply focused on forensically sound storage of bitstreams,
leaving the necessary activities of translating low level storage into
higher level abstractions at the aggregate block (ie. RAID), volume,
and filesystem layers in the domain of analysis tools, as transiently
constructed artifacts. In contrast AFF4 has mechanisms for describing
transformation in a flexible and concise way, allowing users to view
multiple transformations of the same data with virtually zero storage
cost. This mechanism is an important enabler for interoperable
forensic tools. For example, carved files may be described in terms of
their block allocation sequences from an image, rather than requiring
the carved file to be copied again.

This paper extends previous work on the Advanced Forensic Format (AFF)
by taking many of the concepts developed and designing a new
specification and toolset. The AFF4 format is a complete redesign of
the architecture. The new architecture is capable of storing multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. The AFF4 format extends to
become a Global Distributed Evidence Management system, while
maintaing compatibility with the old AFF file format.

We call the new system AFF4, and use the phrase AFF1 to refer to the
legacy system developed by Garfinkel et al.\footnote{Although Garfinkel
never changed the AFF bit-level specification, Garfinkel released
AFFLIB implementations with major version numbers 1, 2 and 3. We
therefore call our system AFF4 to avoid confusion.}

%\begin{figure}
%\fbox{
%\begin{minipage}{.95\linewidth}
%\begin{itemize}
%\item The ability to store evidence from one or more physical
%  devices.
%\item Schema allowing evidence files to contain memory
%  images, network packet intercepts, Netflow data, and extracted
%  files.

%\item Direct encoding of arbitrary mapping transformations of one or
%more sources to allow multiple logical views. This allows 
%RAID volumes to be automatically reassembled from evidence files made
%of the physical RAID devices. Maps can be created using Sleuthkit and
%stored in the evidence file, thereafter allowing arbitrary programs to directly
%read individual files stored in a disk image.

%\item A revised compression system that allows individual chunks
%  within a 16MB segment to be decompressed without decompressing the
%  entire segment.

%\item Support for Global Distributed Evidence Management.
%\end{itemize}
%\end{minipage}}
%\caption{Specific design goals for the next generation forensic file
%  format.}
%\end{figure}

\section{The Need for an Improved Forensic Format}

AFF1's flexibility came from a data model of forensic data and metadata
stored as arbitrary name/value pairs called \emph{segments}. For
example, the first 16MB of a disk image is stored in a segment called
\texttt{page0}, the second 16MB in a segment called \texttt{page1},
\emph{etc.} Because of this flexibility, it was relatively easy 
to extend AFF1 to support encryption, digital signatures, and the
storage of new kinds of metadata such as chain-of-custody
information\cite{garfinkel:affcrypto}.

\subsection{AFF Limitations}
We observed a number of practical problems in the underlying AFF1
standard and Garfinkel's AFFLIB implementation:

@@

\item AFF1 (and the EWF format) are unable to recover from loss or
  corruption. Due to the specialised nature of these formats,
  specialised recovery software must be written to recover from
  corruptions.

@@

\item Although the AFF1 specification calls for a ``table of contents'' similar
  to the Zip\cite{zip-format} ``central directory'' that is stored at
  the end of AFF files, Garfinkel never implemented this directory in
  the publicly released AFF1 implementation, AFFLIB. As a result,
  every header of every segment in an AFF file needs to be read when a
  file is opened. In practice this can take up to 10--30 seconds the
  first time a large AFF file is opened.

%(Once the file is opened, the
%  sectors corresponding to the segment headers are stored in the
%  computer's disk cache.)

@@

\section{Global Distributed Evidence Management}
While AFF1 was designed for use on a single machine that could both
image evidence and perform analysis, many modern practitioners work in
distributed environments in which imaging and analysis takes place in
multiple locations and is performed by multiple individuals.

%This is a
%state of affairs which mirrors common practice in the field.

@@

Managing evidence in a globally distributed system requires the use of
globally unique identifiers to ensure no name collisions can occur
with disconnected systems. AFF1 assigns each piece of evidence a
unique 128-bit identifier called a GID but did not make it clear when
this identifier should be changed and when it should remain the same.

Consider the typical usage scenario depicted in Figure \ref{usage}, of
a volume containing a disk image. This volume is distributed to two
independent analysts, Alice and Bob. Alice may find and extract
individual files, while Bob may correlate information in the evidence
file with other data that is available on departmental
servers. Although in some environments Alice and Bob may be able to
work on a shared file that is located on a server, in other
environments there will not be sufficient connectivity. Instead, each
analyst will be required to store the information in their own
evidence file; these files will then be recombined at a later point in
time.

@@

We have also released a library and toolset, \emph{LIBAFF4}
implementing the new schema and file format. This new library provides
legacy support for existing AFF1 volumes.

@@

This section discusses the AFF4 terminology and architecture. The AFF4
design is object oriented, in that a few generic objects are presented
with externaly accessible behaviour. We discuss a number of
implementations of these high level concepts and show how these can be
put together in common usage cases.

@@

\label{definitions}
\item \emph{A stream} is an AFF Object which provides the ability to
  seek and read random data. Stream objects implement abstracted
  storage, but must provide clients the stream like interface. For
  example, we discuss the {\em Image stream} used to store large
  images, the {\em Map stream} used to create transformations and the
  {\em Encrypted stream} used to provide encryption.

\item \emph{A segment} is a single unit of data written to a volume. AFF4
  segments have a \emph{segment name} provided by their URN, a
  \emph{segment timestamp} in GMT, and the \emph{segment
  contents}. Segments are suitable for storing small quantities of
  data, and still present a stream interface.

@@

\section{Metadata and the Universal Resolver}
\label{resolver}
%One of the fundamental tasks of forensics is in ensuring the
%provenance of any particular piece of digital evidence, for example by
%employing a chain of custody process. To assist in this, acquisition
%methodologies emphasise the process of individuating evidence, and
%storing it in a manner that is easily identifiable.

@@

%In this way we can identify uniquely any evidence without the chance
%for name collisions. Once a universal identification regime is
%created, it is now possible to collect metadata about each object.

The AFF4 model treats metadata as an abstract concept which may exist
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in triple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. Using this system we are able to store
arbitrary attributes about any object in the AFF4
universe. Additionally, as these statements are universally scoped,
they may be stored anywhere.

The AFF4 design extends beyond the management of a single volume,
stream or image to a universal system for managing data of many
types. This necessarily means that a single running instance is
generally unable to have visibility of the entire AFF4 universe. For
example, if a volume is opened which contains a Map Stream targeting a
stream stored in a different volume, it is not generally possible to
tell in which volume the stream exists, and where that volume is
actually stored.

To provide this global visibility of metadata we define a central
metadata management entity, named the {\em Universal Resolver}. The
Universal Resolver contains all the metadata about the AFF4 universe,
that is to say it is able to resolve queries for any attribute about
any URN in the universe.

@@

To this end we define a way for serializing metadata statements (or
triples) into a standard format which implementations can load into
their respective resolvers when parsing the volume. Triples can be
stored in segments having a URN ending with ``{\em properties}''. The
AFF4 implementation loads these segments automatically into the
Universal Resolver.

@@

When the volume is loaded, the AFF4 implementation automatically loads
any properties files and populates its Universal Resolver with the
information visibile to it. AFF4 provides a mechanism to use an
external resolver stored on a MySQL database to provide for a
Universal Resolver that shares information between different instances
on the same network.

@@

\section{Volumes}
The volume object is responsible for providing storage for
segments. Segments are stored and retrieved using their URNs. We
describe two different implementations of volume objects, namely the
{\em Directory Volume} and the {\em ZipFile Volume}. It is possible to
convert from one implementation to another easily, without affecting
any external references. This is because external references are to
the Volume URN itself, so it is safe to convert from one
representation to another. In this way its possible for example, to
unzip a Zip64 volume into a directory and vice versa.

It is important to emphesise that Volumes are merely containers which
provide storage for segments. There is no restriction of which
segments can be stored by any particular volume. For example, the
segments which make up a single Image stream may be stored in a number
of volumes (splitting the image in some way among them). Similarly,
the segments representing a number of streams may be stored in the
same volume.

@@

\begin{itemize}
\item There is already wide support for the Zip and Zip64 formats. By
  migrating to these formats, we can take advantage of the rich number
  of user and developer tools already available. The volume may be
  inspected using any number of commercial or open source zip
  application (e.g. Windows Explorer natively supports Zip files as
  can be seen in Figure~\ref{explorer}, Zip64 is supported natively
  by Java, Python and PERL).

@@

  \caption{The basic structure of a Zip archive. Also shown is how new
  archive members are added to an existing Zip File. The Central
  Directory is overwritten by the new member, and a new Central
  Directory is written on the end.  }

@@

\item Although Zip64 also defines encryption and authentication
extensions, we do not use them due to the restrictions imposed on
their use and because they lack the functionality that is important
for a forensic user. Instead, we use AFF4's digital signature
facilities for integrity and non-repudiation, and we introduce a new
stream based encryption scheme for ensuring data privacy
(Section~\ref{crypted_stream}).

@@

\subsection{Implementing Zip64 in AFFLIB}
Although there are numerous Zip implementations available today, we
have created our own implementation. There are many reasons to develop
our own Zip64 implementation for AFF4:

@@

Storage for the data is done by using multiple data segments stored on
various volumes. Data segment URNs are derived by appending an 8
digit, zero padded decimal integer representation of an incrementing
id to the stream URN (e.g. ``urn:aff4:83a3d6db-85d5/00000032''). Each
data segment is called a \emph{bevy} and stores a number of compressed
chunks back to back.

@@

Image streams specify the \texttt{chunk\_size} attribute, as the
number of image bytes each chunk contains (typically chunk sizes are
32kb). Also specified is the \texttt{chunks\_per\_segment} attribute
which specifies how many chunks are stored in each bevy. Each chunk is
compressed individually using the zlib compress algorithm. This
general structure of storing chunks within larger segments is similar
to the technique used by the Expert Witness file format (EWF) used by
EnCase\cite{encase-3.0} and implemented by the open source
libewf\cite{libewf} package. This improvement from AFF1's 16MB segment
size results in a better match between requested size and the minimum
size required for decompression. This means that less data is needed
to be decompressed unnecessarily where reading small sectors randomly.

@@

The mapping function is described in a segment named by appending
``/map'' to the stream URN. The segment data consists of a series of
lines, each containing a stream offset, a target offset and a target
URN. Offsets are encoded using decimal notation.

@@

The Encrypted Stream may contain any data at all, including disk
images, network captures or memory images. It is useful however, to
store an entire AFF4 volume within the Encrypted stream. This provides
block level encryption for the contained AFF4 volume (which might
contain arbitrary streams). This approach is illustrated in
Figure~\ref{crypted_fif}.

@@

The result is that a number of AFF4 volumes are used as {\em Container
Volumes} to provide storage for Encrypted Streams. The main {\em
Embedded Volume}, which actually contains data is stored within the
Encrypted Stream, effectively distributed throughout the container
volumes. Note that the outer Volume may contain several Encrypted
Streams and therefore contain multiple AFF4 Encrypted
Volumes. Container Volumes may contain non encrypted streams as well,
and may implement different encryption schemes and keys for each
Encrypted stream. This effectively allows arbitrary access policies to
be implemented as only volumes which can be accessed can be merged.

\subsection{The Link Object}
Although the URN of a stream names it unambiguously in the AFF4
universe it is difficult to use and communicate due to its random
nature. Most investigators would prefer to use a shorter name which
might well represent the image better in their minds (e.g. a case
name or warrant number).

A Link is an object which has a single attribute
\emph{aff4:target}. When the Link object is opened, it returns the
object named by this attribute. This allows images with complex names
to be referred to via short, meaningful names. Note that in practice
both Image Streams and Link Objects are created by the imaging tools,
so as far as the operators are concerned they can always refer to the
Image stream via the Link name.

\subsection{Identity Object}
AFF4 defines a \emph{Statement} as a collection of facts, or
(subject, attribute, value) tuples. Listing \ref{statement}
illustrates a collection of facts encoded in the standard AFF4
notation (SHA254 hashes are base64 encoded).

\begin{lstlisting}[
	caption=An Example Statement  (URNs and hashes are
shortened for illustration).,
	label=statement,
	float=tb,
	 ]
urn:aff4:34a62f06/00000000 aff4:sha256=+Xf4i....7rPCgo=
urn:aff4:34a62f06/00000000.idx aff4:sha256=ptV7xOK6....C7R6Xs=
urn:aff4:34a62f06/properties aff4:sha256=yoZ9dtv....eUignYMtk=
urn:aff4:34a62f06 aff4:sha256=udajC5C...BVii7psU=
\end{lstlisting}

The statement expresses a set of attributes of other AFF4 objects, and
in particular the attribute of SHA256 hash is expressed (but other
attributes may also be expressed).

Digital signatures have been used in previous forensic file formats
(such as AFF1) to provide authentication and non-repudiation of
forensic evidence. In essence, a when a person signs an object they
are vouching for its authenticity. Similarly, when a person signs a
\emph{Statement}, they are vouching for its authenticity. This concept
is similar to the Bill of Material (BOM) from AFF1.

An AFF4 Identity object represents an entity, currently described by
way of an X509 certificate. The URN of an identity object is the
certificate's fingerprint, and is therefore unique to the
certificate. Identity objects contains \emph{aff4:statement}
attributes which refer to AFF4 streams containing statements. The
identity object also contains a copy of the certificate used to sign
the statements.

To verify the signatures, the AFF4 library loads the stored
certificate, then checks the signature for each statement. If a
statement is verified (i.e. deemed as correct according to the
identity), the facts within it are checked. Note that it is possible
for multiple identities to sign the same data.

%\subsection{Encryption Schemes}
%Our implementation provides a number of encryption schemes.

%\subsubsection{Null Encryption}
%This scheme ({\em scheme=null}) specifies no encryption at all. This
%may be useful for testing an implementation, but obviously provides
%no real security.

%\subsubsection{AES-SHA-PSK}
%This scheme ({\em scheme=aes-sha-psk}) uses AES for block level
%encryption, with the following SHA based ESSIV scheme, and Pre-Shared
%Key (passphrase).

%When creating a new encryption stream, a master key is generated using
%the first 128 bits of a SHA1 hash of the Pre-Shared-Key appended to a
%64 bit random salt:
%\begin{eqnarray}
%Key_{Master} = \left | SHA1(salt + PSK) \right | _{128}
%\end{eqnarray}

%For each chunk, the chunk IV is obtained by taking the first 128 bits
%of the SHA1 hash of the chunk id encoded as a 32 bit little endian
%integer, followed by the master key:
%\begin{eqnarray}
%IV = \left | SHA1(chunk\_id + Key_{Master}) \right | _{128}
%\end{eqnarray}

%The block is then encrypted or decrypted using AES in CBC mode with
%this IV.

%The salt is specified as a stream attribute named {\em aff4:salt} in
%base64 encoding. Implementations are free to specify any API for
%passing the Pre-Shared Key, for example the PSK can be specified in a
%configuration file, typed in at the terminal, passed in using process
%environment variables, or encoded within the AFF4 URL itself.

@@

%\subsection{In-place encryption of AFF4 volumes}
%With the large size of a typical forensic image it is often difficult
%to have to create a second copy of a volume which needs to be
%encrypted. It is usually more convenient to encrypt the file in place,
%overwriting the plain text data with the equivalent encrypted data.
